---
title: "NEON_data_cleanup"
output: html_document
date: "2026-02-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 Load in Tidyverse and Here library
```{r}
library(tidyverse)
library(here)
```

 Read in the raw data for discharge and check structures of the variables. We also want to see the general layout of the data.  
```{r}
discharge_raw <- read_csv(here("data", "raw", "csd_15_min.csv"))

head(discharge_raw)
str(discharge_raw)
view(discharge_raw)
```

Begin wrangling our data by selecting different variables that could be important for future analysis across our different data tables. We also made sure to try and select variables that would be easy to refine and also variables that may be in common with the variables in different data sets. 
```{r}
discharge <- discharge_raw %>% 
  select(domainID,
         siteID,
         startDateTime,
         endDateTime,
         stageContinuous,
         stageTotalUncert,
         dischargeContinuous,
         dischargeUpperParamUncert,
         dischargeLowerParamUncert,
         dischargeUpperRemnUncert,
         dischargeLowerRemnUncert,
         dischargeCorrectionApplied,
         dischargeFinalQF,
         dischargeFinalQFSciRvw)
```

Write a new csv file for the new data selections we made. Also saved this in the processed data folder. 
```{r}

discharge %>% 
  write_csv(here("data", "processed", "discharge_processed.csv"))

```


Here we begun processing the fish data. First we needed to add in the raw data.

```{r}
fish_raw <- read_csv(here("data", "raw", "fsh_perFish.csv"))

head(fish_raw)
str(fish_raw)
view(fish_raw)
```
```{r}
fish <- fish_raw %>% 
  select(domainID,
         siteID,
         namedLocation,
         passStartTime,
         passEndTime,
         passNumber,
         boutEndDate,
         eventID,
         taxonID,
         taxonRank,
         scientificName,
         morphospeciesID,
         identificationQualifier,
         dataQF,
         specimenNumber,
         fishTotalLength,
         fishWeight,
         samplerType,
         efMortality,
         efInjury,
         fishLifeStage)
```
here we placed fish_processed in the processed data file 
```{r}

fish %>% 
  write_csv(here("data", "processed", "fish_processed.csv"))

```
here we read in the plant_raw data, and narrowed downt he plant variables we will be using to plot.
```{r}
plant_raw <- read_csv(here("data", "raw", "apc_perTaxon.csv"))

head(plant_raw)
str(plant_raw)
view(plant_raw)
```
```{r}
plant <- plant_raw %>% 
  select(domainID,
         siteID,
         namedLocation,
         collectDate,
         sampleCondition,
         aquaticPlantType,
         growthForm,
         taxonID,
         scientificName,
         morphospeciesID,
         phylum,
         division,
         class,
         order,
         family,
         genus,
         section,
         specificEpithet,
         infraspecificEpithet,
         variety,
         form,
         taxonRank,
         identifiedDate,
         dataQF)
```
here we placed plant_processed in the processed data file 
```{r}

plant %>% 
  write_csv(here("data", "processed", "plant_processed.csv"))

```
## Preliminary Plot
```{r}
ggplot(data = discharge,
       aes(x = startDateTime, 
           y = dischargeContinuous,
           col = siteID)) +
  geom_point(cex = 1) +
  stat_smooth()
```


You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
